---
title: "The Effect of A Promotion Program on Sales and Customer Review Attitude For a Book Company"
output: html_document
date: "2023-02-20"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(readr)
library(lubridate)
library(dplyr)
```

## Get familar with the Data

It is a data frame with **5000 observations** of **5 variables**.

There are five columns, date (character), user_submitted_review (character), title (character), total_purchased (integer) and customer type (character).

A brief look at the data:

* date                 : chr  "2019-05-22" "2019-11-16" "2019-06-27" "2019-11-06" ...
* user_submitted_review: chr  "it was okay" "Awesome!" "Awesome!" "Awesome!" ...
* title                : chr  "Secrets Of R For Advanced Students" "R For Dummies" "R For Dummies" "Fundamentals of R For Beginners" ...
* total_purchased      : int  7 3 1 3 NA 1 5 NA 7 1 ...
* customer_type        : chr  "Business" "Business" "Individual" "Individual" ...

There are **two** columns that have NA values.

* **User_submitted_review** column has 456 NA values
* **Total_purchased** column has 718 NA values


```{r importing data, echo=T, results=FALSE}
getwd()
library(tidyverse)
library(readr)
raw<-read.csv("Part 1 sales2019.csv")
str(raw)
table(is.na(raw$date))
table(is.na(raw$user_submitted_review))
table(is.na(raw$title))
table(is.na(raw$total_purchased))
table(is.na(raw$customer_type))

```

## Handling the Missing Data

We're going to remove any rows that have missing data in user_submitted_review and, for the total_purchased column, replace all of the NA values with an average value that we calculate from the complete dataset.

There are **4544 observations** left in the clean data frame.

```{r pressure, echo=TRUE}
fillin<-mean(raw$total_purchased, na.rm = TRUE)
clean <- raw %>%
  filter(!is.na(user_submitted_review)) %>%
  mutate(new_total_purchased = if_else(is.na(total_purchased), fillin, total_purchased))

#print the first few rows to check
head(clean)
table(is.na(clean$new_total_purchased))

```


## Processing Review Data
I want to be able to classify reviews as either **positive, negative or neutral**.
The comments are one of the followings:

* "Awesome!"                              (positive comment)
* "Never read a better book"              (positive comment)
* "I learned a lot"                       (positive comment)

* "it was okay"                           (neutral comment)
* "OK"                                    (neutral comment)
* ""                                      (neutral comment)

* "Hated it"                              (negative comment)
* "The author's other books were better"  (negative comment)
* "A lot of material was not needed"      (negative comment)
* "Would not recommend"                   (negative comment)


Based on the comment, I categorized the comments in a new column called "comment_attitude".

```{r echo=TRUE}
unique(clean$user_submitted_review)
cleanreview <- clean %>% 
  mutate(
    comment_attitude = case_when(
      user_submitted_review %in% c("it was okay", "OK", "") ~ "neutral comment",
      user_submitted_review %in% c("Awesome!", "Never read a better book", "I learned a lot") ~ "positive comment",
      TRUE ~ "negative comment"
    ) 
  )

head(cleanreview)

```

## Comparing Book Sales Between Pre- and Post-Program Sales

The program started on July 1, 2019 and the data I have contains all of the sales for 2019.
Here I will convert the date from string to a date, create a separate column, "before_after", to distinguish the sales made before and after the program, and compare the sum of sales before and after the program was implemented.


```{r echo=TRUE}
#conversion
cleanreview$date<-ymd(cleanreview$date)
str(cleanreview)

#separate column
clean_review_program<-cleanreview %>%
  mutate(before_after = ifelse(cleanreview$date>"2019-07-01","After","Before")
  )

#sum of sales before and after
sales_before_after<-clean_review_program %>% 
  group_by(before_after) %>% 
  summarize(
    books_purchased = sum(new_total_purchased)
  )
sales_before_after
```

There were around **9141 purchases made before the program** and around **9051 made after the program**. Just by this number, it seems that the program even harmed the number of sales a little.

I would like to divide the purchases by customer type in the next step.


## Book Sales Within Customer Type Before and After the Program

Here is the summary of numbers of books purchased before and after program by both business and individuals.

```{r echo=TRUE, message=FALSE, warning=FALSE}
clean_review_program_customer<-clean_review_program %>% 
  group_by(before_after, customer_type) %>% 
  summarize(
    books_purchased = sum(new_total_purchased)
  )
clean_review_program_customer

```

```{r echo=TRUE}
ggplot(clean_review_program_customer, aes(x=customer_type, y=books_purchased, fill=before_after)) +
  geom_bar(stat="identity", position="dodge") +
  scale_fill_manual(values = c("tan1", "steelblue3"))
```


It seems that Business and individuals reacted differently to the program. **Purchases made by business decrease slightly** (from 6300 to 6235) after the program. However, **purchases made by individuals increased** from 2751 to 2906.


## Review Attitude Before and After the Program

```{r echo=TRUE}
review <- clean_review_program %>% 
  group_by(before_after) %>% 
  summarize(
    num_positive_reviews = sum(comment_attitude == "positive comment"),
    num_neutral_review = sum(comment_attitude == "neutral comment"),
    num_negative_review = sum(comment_attitude == "negative comment")
  )
review
```

There is an influence on the customer review attitude by the program. There are **more positive reviews after the program**. However, the difference **might be negligible** until further testing.

## Conclusion

The goal was to find out how effective the program was. After looking at the effect that it had on the sales overall, I believe we can conclude that **it wasn’t an effective program.**
Although there was a slight change between its effect on business customers and individual customers, without m a hypothesis test, I cannot confirm that the difference between business and individual customers' reactions is significant or the difference between the sales data before and after the program to confirm is significant.




